---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
<body style="background-color:LightSkyBlue;">


## Environment setup and configuration
```{r echo=FALSE}
## 
## Reproducible Data - Project 1
##
library(dplyr)
library(ggplot2)
library(knitr)
opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
Code that is needed to load and process the activity data.  

**1. Load the data**  
```{r echo=TRUE}
## Read in raw data from activity.csv file
rawdata <- read.csv("activity.csv",header=TRUE,na.strings="NA")
```

**2. Process/transform the data**  
```{r echo=TRUE}
## Transform the data convert date attribute to valid Date - NEW DATASET
data <- rawdata %>% mutate(date=as.Date(date,format = "%Y-%m-%d"))
```


## What is mean total number of steps taken per day?
The mean total number of steps taken per day, ignoring any missing data in the dataset.  

**1. Calculate the total number of steps taken per day**  
```{r echo=TRUE}
## Data filtered and grouped by date for a summary of steps per day
dailysteps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
    print(dailysteps)

```

**2. Histogram of the total number of steps taken each day**  
```{r echo=TRUE}
# Basic histogram of summarized steps by day
ggplot(dailysteps, aes(x=steps)) + 
    geom_histogram(binwidth=1000, color="blue", fill="white") +
    labs(title = "Histogram of Steps per day", x = "Steps", y = "Frequency") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.5))
```

**3. The the mean and median total number of steps taken per day**  
<p style="color:red;"><b>NOTE:  ignore the missing values in the dataset</b></p> 
```{r echo=TRUE}
## Calculation of mean steps taken each day.
meansteps <- mean(dailysteps$steps, na.rm=TRUE)
    print(meansteps)
## Calculation of median steps taken each day.
medsteps <- median(dailysteps$steps, na.rm=TRUE)
    print(medsteps)
```


## What is the average daily activity pattern?
Summary of the average daily activity pattern.  

**1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**  
```{r echo=TRUE}
## Data filtered and grouped by interval for mean of steps per day
intvlsteps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(interval) %>%
    summarize(steps = mean(steps))
    print(intvlsteps)
## Basic time series of mean steps per day
ggplot(intvlsteps, aes(x=interval, y=steps)) +
    geom_line(color = "blue") +
    labs(title = "Time Series of Steps per Interval", x = "Interval", y = "Steps") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.5))
```

**2. The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps**  
```{r echo=TRUE}
## Data evaluated to determine interval with maximum steps, on average, across all the days
maxintvl <- intvlsteps[which.max(intvlsteps$steps),]
    print(maxintvl)
```


## Inputing missing values
<p style="color:red;"><b>NOTE: There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some   calculations or summaries of the data.</b></p>    

**1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**   
```{r echo=TRUE}
## Summarize total missing values in the dataset
nodata <- sum(is.na(data$steps))
    print(nodata)
```

**2. Devise a strategy for filling in all of the missing values in the dataset. The mean for a specific 5-minute interval used to replace NA values**  
```{r echo=TRUE}
## Data evaluated to calculate the number of steps taken in each 5-minute interval per day
## Transform the data replacing NA steps with the average number of steps in the same 5-min interval
## from the date transformed data, use tapply for filling in the missing values  - NEW DATASET
datafill <- data
nas <- is.na(datafill$steps)
avgintvl <- tapply(datafill$steps, datafill$interval, mean, na.rm=TRUE, simplify=TRUE)
datafill$steps[nas] <- avgintvl[as.character(datafill$interval[nas])]
checkdata <- sum(is.na(datafill$steps))
    print(checkdata)
```

**3. A new dataset that is equal to the original dataset but with the missing data replaced**  
```{r echo=TRUE}
## Data evaluated to calculate the number of steps taken in each 5-minute interval per day
stepsfill <- datafill %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
    print(stepsfill) 
```

**4. Histogram of the total number of steps taken each day and calculation to report the mean and median total number of steps taken per day**    
```{r echo=TRUE}
## Basic histogram of summarized steps by day with NA values replaced with interval average data
ggplot(stepsfill, aes(x = steps)) +
    geom_histogram(binwidth = 1000, color="blue", fill = "white") +
    labs(title = "Histogram of Steps per Day - Missing Data Replaced", x = "Steps", y = "Frequency") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.5))
```


## Are there differences in activity patterns between weekdays and weekends?
<p style="color:red;"><b>NOTE:  Use the dataset with the filled-in missing values for this part.</b></p>    

**1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day**   
```{r echo=TRUE}
## Transform the data replacing NA steps with the average number of steps in the same 5-min interval
## from the date transformed data, use tapply for filling in the missing values  - NEW DATASET
datafill <- data
nas <- is.na(datafill$steps)
avgintvl <- tapply(datafill$steps, datafill$interval, mean, na.rm=TRUE, simplify=TRUE)
datafill$steps[nas] <- avgintvl[as.character(datafill$interval[nas])]
## checkdata <- sum(is.na(datafill$steps))
##    print(checkdata)
## Transform the step NA filled data, add factor variable for weekday or weekend - NEW DATASET
datalevels <- mutate(datafill, weektype = ifelse(weekdays(datafill$date) == "Saturday" | weekdays(datafill$date) == "Sunday", "weekend", "weekday"))
    str(datalevels)
## Data evaluated to calculate the average steps in the 5-minute interval for weekday and weekend
intervallevels <- datalevels %>%
    group_by(interval, weektype) %>%
    summarize(steps = mean(steps))
    print(intervallevels)
```

**2. A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekdays and weekend days (y-axis)**  
```{r echo=TRUE}
## Basic time series of mean steps per day weekday vs weekend
ggplot(intervallevels, aes(x=interval, y=steps, color = weektype)) +
    geom_line() +
    facet_wrap(~weektype, ncol = 1, nrow=2) +
    labs(title = "Time Series of Steps per Interval by Weektype", x = "Interval", y = "Steps") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.33))
```