---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
<body style="background-color:LightSkyBlue;">

## Environment setup and configuration
```{r echo=TRUE}
## 
## Reproducible Data - Project 1
##
library(dplyr)
library(ggplot2)
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

1. Load the data  
```{r echo=TRUE}
## Read in raw data from activity.csv file
rawdata <- read.csv("activity.csv",header=TRUE,na.strings="NA")
```

2. Process/transform the data  
```{r echo=TRUE}
## Transform the data convert date attribute to valid Date - NEW DATASET
data <- rawdata %>% mutate(date=as.Date(date,format = "%Y-%m-%d"))
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day  
```{r echo=TRUE}
## Data filtered and grouped by date for a summary of steps per day
dailysteps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
    print(dailysteps)

```

2. Histogram of the total number of steps taken each day  
```{r echo=TRUE}
# Basic histogram of summarized steps by day
ggplot(dailysteps, aes(x=steps)) + 
    geom_histogram(binwidth=1000, color="blue", fill="white") +
    labs(title = "Histogram of Steps per day", x = "Steps", y = "Frequency") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.5))
```

3. The the mean and median total number of steps taken per day    
    NOTE:  ignore the missing values in the dataset  
```{r echo=TRUE}
## Calculation of mean steps taken each day.
meansteps <- mean(dailysteps$steps, na.rm=TRUE)
    print(meansteps)
## Calculation of median steps taken each day.
medsteps <- median(dailysteps$steps, na.rm=TRUE)
    print(medsteps)
```

## What is the average daily activity pattern?

1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
```{r}

```

2. The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps  
```{r}

```

## Inputing missing values
Note: There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)  
```{r}

```

2. Devise a strategy for filling in all of the missing values in the dataset. The mean for a specific 5-minute interval used to replace NA values.
```{r}

```

3. A new dataset that is equal to the original dataset but with the missing data replaced. 
```{r}

```

4. Histogram of the total number of steps taken each day and calculation to report the mean and median total number of steps taken per day.   
```{r}

```

## Are there differences in activity patterns between weekdays and weekends?


```{r}

```