---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
<body style="background-color:LightSkyBlue;">

## Environment setup and configuration
```{r echo=TRUE}
## 
## Reproducible Data - Project 1
##
library(dplyr)
library(ggplot2)
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

1. Load the data
```{r echo=TRUE}
## Read in raw data from activity.csv file
rawdata <- read.csv("activity.csv",header=TRUE,na.strings="NA")
```

2. Process/transform the data
```{r echo=TRUE}
## Transform the data convert date attribute to valid Date - NEW DATASET
data <- rawdata %>% mutate(date=as.Date(date,format = "%Y-%m-%d"))
```


## What is mean total number of steps taken per day?
```{r echo=TRUE}
## Data filtered and grouped by date for a summary of steps per day
dailysteps <- data %>%
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
    print(dailysteps)

```

1. Histogram of the total number of steps taken each day
```{r echo=TRUE}
# Basic histogram of summarized steps by day
ggplot(dailysteps, aes(x=steps)) + 
    geom_histogram(binwidth=1000, color="blue", fill="white") +
    labs(title = "Histogram of Steps per day", x = "Steps", y = "Frequency") +
    theme(plot.title.position = 'plot',plot.title = element_text(hjust = 0.5))
```

2. The the mean and median total number of steps taken per day
```{r echo=TRUE}
## Calculation of mean steps taken each day.
meansteps <- mean(dailysteps$steps, na.rm=TRUE)
    print(meansteps)
## Calculation of median steps taken each day.
medsteps <- median(dailysteps$steps, na.rm=TRUE)
    print(medsteps)
```

## What is the average daily activity pattern?

<p style="color:red"><b>Here is some text, this part of which is red!</b></p>

## Inputing missing values
```{r}

```



## Are there differences in activity patterns between weekdays and weekends?
